A-maze of Natural Stories: Texts are comprehensible using the Maze task

### 2020-04-30 Papers/amlap2020_abstract.tex
revisions to abstract

\title{A-maze of Natural Stories: Texts are comprehensible using the Maze task}
\authors{Veronica Boyce (MIT), Roger Levy (MIT)} %TODO, check what affiliation I have at conference time
\email{vboyce@mit.edu}

% Here goes the main text of your abstract:
%Top overview
A-maze is a promising new method for measuring incremental sentence processing (Boyce et al. 2020; Sloggett et al. 2020). However, so far A-maze has only been used on constructed psycholinguistic items, so it is unknown how fluently people can comprehend and remember what they read during the Maze task. %TODO do I need to motivate why people might not comprehend?
%TODO do I need to introduce what the Maze task is earlier?
We address this question by running A-maze on the stories from the Natural Stories Corpus (Futrell et al. 2017). We find that some participants can follow the stories and correctly answer comprehension questions even when reading with Maze. We also find a linear effect of a word's surprisal on reaction time, but little effect of the surprisal of the previous word, indicating a lack of spill-over effects.

% How does A-maze work 
In the Maze task, participants read a sentence word by word (see Figure~1). For each word position, they see two words, one of which is the next word in the sentence and one of which is a distractor. Participants press a key to indicate which word continues the sentence; the time between key presses (reaction time, or RT) is the dependent measure.  Usually in the Maze task, when a participant makes a mistake, the sentence stops and they move on to the next item. In order to present coherent texts, we instead have participants correct their mistakes (as shown in Figure~1). When a participant makes a mistake, they see an error message and must press the correct key to continue with the sentence. This way, participants see all the content of each sentence, and so can follow the story as it continues in the next sentence.

 The Natural Stories Corpus contains 10 passages, each about 1000 words long and 6 comprehension questions per passage (Futrell et al. 2017). We used the A-maze framework from Boyce et al. (2020) and the language model from Gulordava et al. (2017) to generate distractor words for the texts. We recruited 100 participants from Amazon Mechanical Turk; each participant did a short practice passage to become familiar with the Maze task, followed by reading one of the Natural Stories in the Maze task, and answering the corresponding comprehension questions.

 We excluded 5 participants who did not report being native English speakers. Some participants appear to be randomly pressing keys to get through the task quickly, while others seem to do the task more slowly and diligently (Figure~2). Task accuracy is correlated with performance on the comprehension questions ($r^2=.47$); of the 63 participants who had at least 80\% accuracy on the Maze task, 50 got 5 or 6 of the 6 comprehension questions correct (Figure~4). We exclude participants with less than 80\% accuracy from analyses.

We wanted to confirm whether a word's surprisal, unigram frequency, and length are predictive of RT. We get surprisal estimates from  3 models: a Kneser-Ney smoothed 5-gram model, a recurrent neural network model (GRNN) from Gulordava et al. (2018), and a transformer model (TXL) from Dai et al. (2019). We fit BRMS models using surprisal, frequency, length, and frequency:length interaction of the word and the previous word as predictors of RT.  We only include words that occurred before any mistakes in their sentence. We only analyse words that the models treat as single tokens.  Results are shown in Table 1; across all models, surprisal and length of the current word show sizeable effects, and properties of the previous word show smaller effects, suggesting a lack of spillover. To confirm the shape of relationship between surprisal and RT, we fit GAMs; as shown in Figure 3, the relationship between surprisal and RT is roughly linear at the current word, and there is little effect for the prior word's surprisal.  %TODO this result also holds? if we include RT's from after a mistake

%TODO am I allowed to use the word surprisal?
% Take away
Our results provide evidence that the A-maze task can be used effectively on naturalistic texts where comprehension is important. We present a variation on the Maze task, where participants correct their mistakes and continue reading, so they don't miss parts of the story. This also makes it easy to exclude participants who are not doing the task attentively. We find that attentive participants can generally follow the story they are reading, and we confirm that A-maze RTs show a strong relationship with the surprisal of the current, but not the past, word. We make our code and data available at [link to repo redacted for anonymity].

@@

		\end{small}
	\end{minipage}
\end{figure}
\vspace{1em}
\begin{figure}
		\begin{minipage}{.5\textwidth}
		\captionof{figure}{Surprisal and RT\\
		}
		{\center\includegraphics[width=\textwidth]{gam.pdf}\\} 
		\begin{small}
			Current word surprisal is linearly predictive of RT, but previous word surprisal is not predictive of RT.

		\end{small}
	\end{minipage}	
	~~~
\begin{minipage}{.5\textwidth}
	\captionof{figure}{Comprehension question accuracy}
	{\center\includegraphics[width=\textwidth]{comp.pdf}\\} 
	\begin{small}
		Participants with higher accuracy on the Maze task also did better on comprehension questions.

@@

	\setlength{\tabcolsep}{4pt}
		\captionof{table}{Regression Coefficients}
\begin{small}%TODO do I even include p-values??
%\centering
\begin{tabular}{l|rlr|rlr|rlr}
	\hline
	&\multicolumn{3}{c|}{5-gram}&\multicolumn{3}{c|}{GRNN}&\multicolumn{3}{c}{TXL}\\
		 & Est & CI & $p$ & Est & CI & $p$ &Est & CI & $p$ \\ 
		 \hline
Intercept & 867.7 & [829.6, 904.3] & 0.00 & 868.5 & [835.4, 903] & 0.00 & 866.5 & [830.4, 900.7] & 0.00 \\
Surprisal & 10.6 & [8.6, 12.6] & 0.00 & 22.1 & [19.8, 24.4] & 0.00 & 17.0 & [14.8, 19.2] & 0.00 \\
Frequency & -4.2 & [-7.3, -0.9] & 0.01 & 1.7 & [-1.2, 4.6] & 0.25 & -0.8 & [-3.7, 2.2] & 0.59 \\ 
Length & 20.1 & [15.3, 24.7] & 0.00 & 18.0 & [13.6, 22.4] & 0.00 & 20.8 & [16.4, 25.3] & 0.00 \\ 
Freq : Length & 1.1 & [0.4, 1.8] & 0.00 & 1.2 & [0.4, 1.9] & 0.00 & 1.2 & [0.5, 2] & 0.00 \\ 
\hline
Past Surprisal & 1.9 & [0.3, 3.4] & 0.02 & 1.9 & [0.6, 3.3] & 0.01 & 0.1 & [-1.2, 1.3] & 0.95 \\ 
Past Freq & 3.0 & [0.8, 5.2] & 0.01 & 1.5 & [-0.2, 3.3] & 0.09 & 0.9 & [-0.9, 2.8] & 0.35 \\ 
Past Length & -5.0 & [-8.7, -1.5] & 0.01 & -6.9 & [-10.5, -3.3] & 0.00 & -5.3 & [-9, -1.7] & 0.00 \\ 
Past Freq : Length& -1.0 & [-1.7, -0.3] & 0.01 & -1.3 & [-1.9, -0.6] & 0.00 & -1.2 & [-1.9, -0.5] & 0.00 \\ 
\hline
\end{tabular}
\vspace{.5em}

\end{small}
		\begin{small}
			Point estimates, credible intervals, and p-value equivalents. Models were run with centered predictors and full mixed effects by subject. Surprisal is in bits, frequency is in $log_2$ occurances per billion words. 
		\end{small}
\vspace{.5em}

	\begin{small}{
			\noindent\textbf{References:}
			Boyce et al (2020). \textit{J Mem. Lang.} $\bullet$
			Dai et al (2019) arXiv:1901.02860 $\bullet$
			Forster et al (2009). \textit{Behav. Res. Methods} $\bullet$
			Futrell et al (2017). arXiv:1708.05763 $\bullet$
			Gulordava et al (2018).  \textit{NAACL-HLT 2018} $\bullet$
			Sloggett et al (2020). \textit{CUNY 2020} $\bullet$
			Witzel et al (2012).  \textit{J Psycholinguist Res} 
			%TODO add cites for length, freq, surprisal justification
		}
	\end{small}

>>>>

\title{A-maze of Natural Stories: Texts are comprehensible using the Maze task}
\authors{Veronica Boyce (MIT), Roger Levy (MIT)} %TODO, check what affiliation I have at conference time
\email{vboyce@mit.edu}
\newline
%
% Here goes the main text of your abstract:
%Comment: I feel like you could make a bigger deal of the current-word previous-word / RT correlation. Given how there are spillover effects for SPR and also for eye-tracking (if I’m not mistaken), it’s crazy that the relationship between RT/Surprisal for the previous word is basically flat, when it comes to Maze!
%Comment: I’m confused with Table 1: Doesn’t length have a bigger effect size than surprisal? On a related note: does it matter which has a bigger effect? Why is it important to say that surprisal has a bigger effect, compared to just a robust, significant effect?

A-maze is a new method for measuring incremental sentence processing that can localize slowdowns related to syntactic ambiguities (Boyce et al. 2020; Sloggett et al. 2020). We test A-maze on the Natural Stories corpus (Futrell et al. 2017) and find that people can comprehend what they read during the Maze task. Moreover, the Maze task yields useable reaction time data with word predictability effects that are linear in the surprisal of the current word, with little spillover effect from the surprisal of the previous word.

The Maze task (Forster et al. 2009; Witzel et al. 2012) is an incremental processing method where participants read a sentence word by word  (Figure~1). For each word position, participants see two words, one of which is the next word in the sentence and one of which is a distractor. Participants press a key to indicate which word continues the sentence; the time between key presses (reaction time, or RT) is the dependent measure.  Traditionally, when a participant makes a mistake, the sentence stops and they move on to the next item. In order to present coherent texts, we instead have participants correct their mistakes (Figure~1). When a participant makes a mistake, they see an error message and must press the correct key to continue with the sentence. This way, participants see all the content and can follow the story, allowing us to test multi-sentence materials.

The Natural Stories Corpus consists of 10 passages ($\sim$1000 words each) with 6 comprehension questions per passage (Futrell et al. 2017). We used the A-maze framework from Boyce et al. (2020) and the language model from Gulordava et al. (2018) to generate distractor words for the texts. We recruited 100 participants from Amazon Mechanical Turk; each participant read one passage in the Maze task and answered the corresponding comprehension questions.

We exclude 5 participants who do not report being native English speakers. Participants clustered into those who appear to randomly press keys to get through the task quickly and those who do the task more slowly and accurately (Figure~2); we exclude participants with less than 80\% accuracy from analyses. Task accuracy is correlated with performance on the comprehension questions ($r^2=.47$); of the 63 participants who had at least 80\% accuracy on the Maze task, 50 got 5 or 6 of the 6 comprehension questions correct (Figure~4). People can understand and remember what they read while doing the Maze task successfully.

We also investigated the relationship between a word's predictability and its RT. We fit Bayesian regression models using suprisal, frequency, length, and surprisal:length and frequency:length interactions at both the current and previous word as predictors of RT (see Table~1 for details). We use 3 models for surprisals: a Kneser-Ney smoothed 5-gram model, a recurrent neural network model (GRNN) from Gulordava et al. (2018), and a transformer model (TXL) from Dai et al. (2019). Across all models, we see large effects of current word surprisal and length, which noticeably exceed effects of previous word predictors (Table~1). We also fit GAMs to look at the shape of the relationship between surprisal and RT; we find a linear relationship between RT and current word surprisal and no relationship with previous word surprisal (Figure~3). The linear relationship matches known effects from eye-tracking and self-paced reading (Smith \& Levy 2013), but the lack of spillover indicates a potential advantage of Maze over eye-tracking and self-paced reading.

In summary, we provide an adaptation of the Maze task for long naturalistic texts, and show that participants can comprehend what they read during A-maze. Additionally, we find robust, localized surprisal effects, supporting Forster et al.'s (2009) argument that Maze forces highly incremental processing. Overall, this suggests that A-maze is a versatile alternative to eye-tracking and self-paced reading.

Our code and data are available at [link to repo redacted for anonymity].

%It’s worth thinking about “why should AMLaP attendees care about this work?” and building an answer to that into the abstract.  Minimally, the work shows
%
%- Maze is really a flexible method
%- you can have people read naturalistic text passages with Maze, they can understand the passages and you wind up with usable data
%- word predictability effects in Maze are linear in surprisal, as they are with SPR and eye tracking
%- potentially unlike SPR or eye tracking, surprisal seems to be far more important than word length or word frequency as a determinant of processing time in Maze (see comments/questions below about the model)
%- also potentially unlike SPR or eye tracking, there isn’t much spillover in Maze for word predictability effects (though there is a little for word length).  All this seems to support Forster et al. 2009’s argument that Maze “forces highly incremental processing”.

@@

		\end{small}
	\end{minipage}
\end{figure}
\begin{figure}
		\begin{minipage}{.5\textwidth}
		\captionof{figure}{Surprisal and RT\\
		}
		{\center\includegraphics[width=\textwidth]{gam.pdf}\\} 
		\begin{small}
			Current word surprisal is linearly predictive of RT, but previous word surprisal is not predictive of RT.

		\end{small}
	\end{minipage}	
	~~~
\begin{minipage}{.5\textwidth}
	\captionof{figure}{Comprehension question accuracy}
	{\center\includegraphics[width=\textwidth]{comp.pdf}\\} 
	\begin{small}
		Participants with high accuracy on the Maze task also perform well on comprehension questions.

@@

	\setlength{\tabcolsep}{3pt}
		\captionof{table}{Regression Coefficients}
\begin{small}%TODO do I even include p-values??
%\centering
\begin{tabular}{l|rlr|rlr|rlr}
	\hline
	&\multicolumn{3}{c|}{5-gram}&\multicolumn{3}{c|}{GRNN}&\multicolumn{3}{c}{TXL}\\
		 & Est & CI & $p$ & Est & CI & $p$ &Est & CI & $p$ \\ 
		 \hline
Intercept & 865.3 & [829.9, 902.9] & 0.00 & 871.1 & [837.9, 905.3] & 0.00 & 870.8 & [832.5, 907.8] & 0.00 \\ 
Surprisal & 11.7 & [9.3, 14.1] & 0.00 & 23.7 & [21, 26.5] & 0.00 & 18.5 & [16.1, 21.1] & 0.00 \\ 
Frequency & -2.9 & [-6.3, 0.5] & 0.10 & 2.9 & [-0.2, 6] & 0.06 & 0.4 & [-2.7, 3.5] & 0.79 \\ 
		Length & 20.5 & [15.4, 25.6] & 0.00 & 18.5 & [13.3, 23.7] & 0.00 & 21.4 & [16.2, 26.6] & 0.00 \\ 
		Surprisal:Length & -2.0 & [-3, -1] & 0.00 & -1.8 & [-2.7, -0.9] & 0.00 & -1.4 & [-2.2, -0.6] & 0.00 \\
		Freq:Length & -1.0 & [-2.5, 0.4] & 0.16 & -0.1 & [-1.2, 1] & 0.82 & 0.2 & [-0.9, 1.2] & 0.76 \\ 
		\hline
			Past Surprisal & 1.6 & [-0.5, 3.6] & 0.14 & 2.7 & [0.8, 4.5] & 0.00 & 0.8 & [-0.9, 2.5] & 0.40 \\ 
		Past Freq & 2.6 & [-0.1, 5.4] & 0.06 & 1.9 & [-0.2, 4.2] & 0.08 & 1.2 & [-1.1, 3.6] & 0.30 \\ 
		Past Length & -4.8 & [-9, -0.1] & 0.04 & -6.6 & [-10.9, -2.1] & 0.00 & -5.2 & [-9.3, -0.7] & 0.03 \\ 
				Past Surp:Length & -0.2 & [-1.2, 0.8] & 0.72 & -0.9 & [-1.7, -0.2] & 0.01 & -0.6 & [-1.3, 0.2] & 0.13 \\ 
		Past Freq:Length & -1.0 & [-2.3, 0.3] & 0.15 & -1.8 & [-2.9, -0.8] & 0.00 & -1.5 & [-2.6, -0.5] & 0.01 \\

\hline
\end{tabular}
\vspace{.1em}

\end{small}
		\begin{small}
			Point estimates, credible intervals, and p-value equivalents. Surprisal was measured in bits, frequency in $log_2$ occurrences per billion words, and length in characters. All predictors were centered, and only single token words were included. Models were fit in BRMS with formula \textit{rt $\sim$ surprisal*length + freq*length + past\_surp*past\_length + past\_freq*past\_length +
			(surprisal*length + freq*length + past\_surp*past\_length + past\_freq*past\_length | participant)+(1|word\_id)}. Results shown for regression on pre-error data only.
		\end{small}

\vspace{.1em}
	\begin{small}{
			\noindent\textbf{References:}
			Boyce et al (2020). \textit{J Mem. Lang.} $\bullet$
			Dai et al (2019) arXiv:1901.02860 $\bullet$
			Forster et al (2009). \textit{Behav. Res. Methods} $\bullet$
			Futrell et al (2017). arXiv:1708.05763 $\bullet$
			Gulordava et al (2018).  \textit{NAACL-HLT 2018} $\bullet$
			Sloggett et al (2020). \textit{CUNY 2020} $\bullet$
			Smith \& Levy (2013). \textit{Cognition} $\bullet$
			Witzel et al (2012).  \textit{J Psycholinguist Res} 
			%TODO add cites for length, freq, surprisal justification
		}
	\end{small}
